**NAME OF THIS STUDY**

### 2018-10-22 submission/manuscript.Rmd
working on ML manuscript

# We want the diagnosis column to a factor
data$dx <- factor(data$dx, labels=c("normal", "cancer"))
```

```{r BinaryClassification, eval=TRUE, echo=FALSE, cache=FALSE, warning=FALSE, message=FALSE}
cl <- makePSOCKcluster(4)
registerDoParallel(cl)
# Create
all.test.response <- all.test.predictor <- test_aucs <- c()
for (i in 1:50) {
  inTraining <- createDataPartition(data$dx, p = .80, list = FALSE)
  training <- data[ inTraining,]
  testing  <- data[-inTraining,]
  x_train <- training %>% select(-dx)
  y_train <- training$dx
  y_train <- as.factor(y_train)
  # L2 regularization is used for Logistic regression
  # cost is the inverse of penalty and we provide a grid
  grid <-  expand.grid(cost = c(0.0000001, 0.000001, 0.00001, 0.0001),
                       loss = "L2_dual",
                       epsilon = 0.1)
  cv <- trainControl(method="repeatedcv",
                     repeats = 50,
                     number=5,
                     returnResamp="final",
                     classProbs=TRUE,
                     summaryFunction=twoClassSummary,
                     indexFinal=NULL,
                     preProc = "scale",
                     savePredictions = TRUE)

  L2Logit <- train(x_train, y_train,
                               method = "regLogistic",
                               trControl = cv,
                               metric = "ROC",
                               tuneGrid = grid,
                               preProcess = c("center", "scale"), 
                               family = "binomial")
  # Mean AUC value of the best lambda parameter training over repeats
  cv_auc <- getTrainPerf(L2Logit)$TrainROC
  # Predict on the test set and get predicted probabilities
  probs <- predict(L2Logit, testing, type="prob")
  # Test AUC calculation
  test_roc <- roc(ifelse(testing$dx == "cancer", 1, 0), probs[[2]])
  test_auc <- test_roc$auc
  # Save all the test AUCs over iterations in test_aucs
  test_aucs <- c(test_aucs, test_auc)
  # Save the test set labels in all.test.response. Labels converted to 0 for normal and 1 for cancer
  all.test.response <- c(all.test.response, ifelse(testing$dx == "cancer", 1, 0))
  # Save the test set predicted probabilities of highest class in all.test.predictor
  all.test.predictor <- c(all.test.predictor, probs[[2]])
}
stopCluster(cl)
# Get the ROC curve and AUC values for all the iterations 
test_roc <- roc(all.test.response, all.test.predictor, auc=TRUE, ci=TRUE)
```

```{r LogitROCplot, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}

pdf("../results/figures/LogReg_inR.pdf")
par(mar=c(4,4,1,1))
# Plot random line on ROC curve
plot(c(1,0),c(0,1),
     type='l',
     lty=3,
     xlim=c(1.01,0), ylim=c(-0.01,1.01),
     xaxs='i', yaxs='i',
     ylab='', xlab='')
# Plot Test ROC in red line
plot(test_roc,
     col='black',
     lwd=2,
     add=T,
     lty=1)
#Calculate Confidence Interval sensitivities at given specificities
sens.ci <- ci.se(test_roc)
# Compute the CI of the AUC
auc.ci <- ci.auc(test_roc)
# Plot CV ROC in blue line
#plot(cv_roc,
#     col='blue',
#     lwd=2,
#     add=T,
#     lty=1)
# Label the axes
mtext(side=2,
      text="Sensitivity",
      line=2.5,
      cex=1.5)
mtext(side=1,
      text="Specificity",
      line=2.5,
      cex=1.5)
# Add legends for both lines
legend(x=0.7,y=0.2,
       legend=(sprintf('Test AUC: %.3g, CI: %.3g', test_roc$auc, auc.ci)),
       bty='n',
       xjust=0,
       lty=c(1,1),
       col='black',
       text.col='black')

>>>>

# We want the diagnosis column to a factor
data$dx <- factor(data$dx, labels=c("normal", "cancer"))
# calculate the pre-process parameters from the dataset
# in python we use StandardScaler, here we center and scale to get the same effect
preprocessParams <- preProcess(data[,-1], method=c("center", "scale"))
# transform the dataset using the parameters
data[,-1] <- predict(preprocessParams, data[,-1])
```

```{r BinaryClassification, eval=TRUE, echo=FALSE, cache=FALSE, warning=FALSE, message=FALSE}
# Create lists to store the predicted labels and actuals labels at each iteration 
all.test.response <- all.test.predictor <- test_aucs <- c()

## We will split the dataset 80%-20% and tune hyper-parameter on the 80% training. This tuning step on training data will be done 50 times with 5 folds on the training set and an optimal hyper-parameter/optimal model will be chosen.

## The chosen best model and hyper-parameter will be tested on the %20 test set that was not seen before during training. This will give a TEST AUC.

## We will do the previous steps 50 iterations. Each iteration will have a unique hyper-parameter and optimal model. For each of these models we will have a TEST AUC. We will report the TEST AUC with 95 %CI.

# For each epoch, we can also keep track of AUC values with CI for each cross-validation during training.But these code lines are not included in this Rmd file. I will have those commented out in the main.R code file. 
for (i in 1:50) {
  # Partition the data 80-20. This function in caret keeps it stratified if I remember correctly.
  inTraining <- createDataPartition(data$dx, p = .80, list = FALSE)
  training <- data[ inTraining,]
  testing  <- data[-inTraining,]
  x_train <- training %>% select(-dx)
  y_train <- training$dx
  y_train <- as.factor(y_train)
  # L2 regularization is used for Logistic regression
  # Cost is our regularization parameter which here is the inverse of penalty and we provide a grid to tune the best one. 
  grid <-  expand.grid(cost = c(0.0000001, 0.000001, 0.00001, 0.0001),
                       loss = "L2_dual",
                       epsilon = 0.1)

  cv <- trainControl(method="repeatedcv",
                     repeats = 50,
                     number=5,
                     returnResamp="final",
                     classProbs=TRUE,
                     summaryFunction=twoClassSummary,
                     indexFinal=NULL,
                     savePredictions = TRUE)

  L2Logit <- train(x_train, 
                   y_train,
                   method = "regLogistic",
                   trControl = cv, 
                   metric = "ROC",
                   tuneGrid = grid,
                   family = "binomial")
  # Mean AUC value of the best lambda parameter training over repeats
  cv_auc <- getTrainPerf(L2Logit)$TrainROC
  # Predict on the test set and get predicted probabilities
  probs <- predict(L2Logit, testing, type="prob")
  # Test AUC calculation
  test_roc <- roc(ifelse(testing$dx == "cancer", 1, 0), probs[[2]])
  test_auc <- test_roc$auc
  # Save all the test AUCs over iterations in test_aucs
  test_aucs <- c(test_aucs, test_auc)
  # Save the test set labels in all.test.response. Labels converted to 0 for normal and 1 for cancer
  all.test.response <- c(all.test.response, ifelse(testing$dx == "cancer", 1, 0))
  # Save the test set predicted probabilities of highest class in all.test.predictor
  all.test.predictor <- c(all.test.predictor, probs[[2]])
}
# Get the ROC curve and AUC values for all the iterations
# auc= TRUE and ci=TRUE will keep track of the auc and 95 CI for all the sensitivities and specificities. 
test_roc <- roc(all.test.response, all.test.predictor, auc=TRUE, ci=TRUE)
```

```{r LogitROCplot, eval=TRUE, echo=FALSE, cache=FALSE, warning=FALSE, message=FALSE}
# Here we plot the test AUC and 95%CI for logistic regression.

pdf("../results/figures/LogReg_inR.pdf")
par(mar=c(4,4,1,1))
# Plot random line on ROC curve
plot(c(1,0),c(0,1),
     type='l',
     lty=3,
     xlim=c(1.01,0), ylim=c(-0.01,1.01),
     xaxs='i', yaxs='i',
     ylab='', xlab='')
# Plot Test ROC in black line
plot(test_roc,
     col='black',
     lwd=2,
     add=T,
     lty=1)
#Calculate Confidence Interval sensitivities at given specificities
sens.ci <- ci.se(test_roc)
# Calculate CI for AUC values
auc_ci <- test_roc$ci[3]-test_roc$ci[2]

# Plot CV ROC in blue line
#plot(cv_roc,
#     col='blue',
#     lwd=2,
#     add=T,
#     lty=1)
# Label the axes
mtext(side=2,
      text="Sensitivity",
      line=2.5,
      cex=1.5)
mtext(side=1,
      text="Specificity",
      line=2.5,
      cex=1.5)
# Add legends for both lines
legend(x=0.7,y=0.2,
       legend=(sprintf('Test AUC: %.3g, %s CI: %0.3g', test_roc$auc, '%95', auc_ci)),
       bty='n',
       xjust=0,
       lty=c(1,1),
       col='black',
       text.col='black')
